Overview

### 2023-11-08 analysis/001_overview.Rmd
Add ta/no ta boxplot in overview

---
title: "Exploratory Data Analysis"
output: word_document
date: "2023-09-11"
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "99%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r dependencies}
library(tidyverse, warn.conflicts = FALSE)
library(cowplot)
```

```{r data}
library(hoaddata)

# ta_oa_inst <- readr::read_csv("https://storage.googleapis.com/hoaddata/ta_oa_inst.csv.gz")
ta_oa_inst <- readr::read_csv(here::here("data/ta_oa_inst.csv.gz"))
```

@@

```{r}
ta_oa_growth <- ta_oa_inst |>
  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424") |>
  group_by(cr_year) |>
  summarise(ta_oa = n_distinct(doi))
hybrid_oa_growth <- hoaddata::cc_articles |>
  filter(issn_l != "0027-8424") |>
  group_by(cr_year) |>
  summarise(oa_all = n_distinct(doi))
ta_df <- inner_join(ta_oa_growth, hybrid_oa_growth) |>
  mutate(ta_oa_prop = ta_oa / oa_all) |>
  mutate(no_ta = oa_all - ta_oa, 
         no_ta_prop = 1 - ta_oa_prop) |>
  mutate(cr_year = as.factor(cr_year))
# year all
year_all <-
  hoaddata::jn_ind |>
  filter(issn_l != "0027-8424") |>
  distinct(issn_l, cr_year, jn_all) |> 
  group_by(cr_year) |> 
  summarise(n = sum(jn_all))

@@

```{r}
ta_share_plot <- ta_df |>
  select(cr_year, ta_oa_prop, no_ta_prop) |>
  pivot_longer(cols = c(ta_oa_prop, no_ta_prop), names_to = "type", values_to = "value") |>
  filter(cr_year != "2023", type == "ta_oa_prop") |>
  ggplot(aes(factor(cr_year), value, group = 1)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25, 0.5),
                     labels = scales::percent) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = "Publication Year")
```

@@

```{r}
library(patchwork)
design = "AC
BC"
wrap_plots(
  A = oa_ta_plot,
  B = ta_share_plot,
  C = oa_jn_box,
  design = design,
  widths = c(2, 1),
  heights = c(2, 1),
  guides = "collect"
) &
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')
```

>>>>

---
title: "Overview"
output: github_document
date: "`r Sys.Date()`"
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "99%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r dependencies}
library(tidyverse, warn.conflicts = FALSE)
```

```{r data, cache=TRUE}
library(hoaddata)

ta_oa_inst <- readr::read_csv("https://github.com/subugoe/hoaddata/releases/download/v0.2.82/ta_oa_inst.csv.gz")
```

@@

```{r area_oa_over_the_years}
ta_oa_growth <- ta_oa_inst |>
  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424") |>
  group_by(cr_year) |>
  summarise(ta_oa = n_distinct(doi))
hybrid_oa_growth <- hoaddata::cc_articles |>
  filter(issn_l != "0027-8424") |>
  group_by(cr_year) |>
  summarise(oa_all = n_distinct(doi))
ta_df <- inner_join(ta_oa_growth, hybrid_oa_growth) |>
  mutate(ta_oa_prop = ta_oa / oa_all) |>
  mutate(no_ta = oa_all - ta_oa, 
         no_ta_prop = 1 - ta_oa_prop) |>
  mutate(cr_year = as.factor(cr_year))
# year all
year_all <-
  hoaddata::jn_ind |>
  filter(issn_l != "0027-8424") |>
  distinct(issn_l, cr_year, jn_all) |> 
  group_by(cr_year) |> 
  summarise(n = sum(jn_all))

@@

```{r ta_share_plot}
ta_share_plot <- ta_df |>
  select(cr_year, ta_oa_prop, no_ta_prop) |>
  pivot_longer(cols = c(ta_oa_prop, no_ta_prop), names_to = "type", values_to = "value") |>
  filter(cr_year != "2023", type == "ta_oa_prop") |>
  ggplot(aes(factor(cr_year), value, group = 1)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25, 0.5),
                     labels = scales::percent) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = "Publication Year")
```

@@

```{r}
ta_oa_growth <- ta_oa_inst |>
  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424") |>
  group_by(cr_year, issn_l) |>
  summarise(ta_oa = n_distinct(doi))
hybrid_oa_growth <- hoaddata::cc_articles |>
  filter(issn_l != "0027-8424") |>
  group_by(cr_year, issn_l) |>
  summarise(oa_all = n_distinct(doi))
ta_df <- left_join(hybrid_oa_growth, ta_oa_growth) |>
  mutate(across(everything(), .fns = ~replace_na(.,0))) |>
  mutate(no_ta = oa_all - ta_oa) |>
  mutate(cr_year = as.factor(cr_year))

summary_all <- inner_join(ta_df, jn_year_all, by = c("cr_year", "issn_l")) |>
  mutate(prop_ta_oa = ta_oa / n,
         prop_no_ta = no_ta / n) |>
  ungroup() |>
  distinct(cr_year, issn_l, prop_ta_oa, prop_no_ta) |>
  pivot_longer(cols = c(prop_ta_oa, prop_no_ta))
# plot
box_ta_prop <- summary_all |>
  filter(cr_year != "2023") |>
  # With at least one OA article
  filter(value > 0) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, fill = name)) +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  coord_cartesian(ylim = c(0,0.5)) +
  scale_fill_manual("Hybrid OA", values = c(prop_no_ta = "#b3b3b3a0", prop_ta_oa = "#56B4E9"), labels = c(prop_no_ta = "Other", prop_ta_oa = "Agreement")) +
  theme_minimal() +
  labs(y = "OA Uptake\nper Hybrid Journal", x = "Publication Year")

```

```{r}
library(patchwork)
design = "AC
BD"
wrap_plots(
  A = oa_ta_plot,
  B = ta_share_plot,
  C = oa_jn_box,
  D = box_ta_prop,
  design = design,
  widths = c(2, 1),
  heights = c(2, 1),
  guides = "collect"
) &
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')
```
